### 2024-02-05 main.tex
20240205

\documentclass{elsarticle}
\usepackage{amsmath,amssymb,amsfonts,amsthm,bm}
\usepackage{hyperref}
\usepackage{booktabs,multirow}
\usepackage{lineno}
\linenumbers

\begin{document}
\include{mix}
\end{document}