### 2024-02-10 introduction.tex
temp

\section{Introduction}
Thin shell is one of the most frequently used structure in engineering practice, where the thickness of this kind structure is often much smaller than its radius. With the Kirchhoff-Love hypothesis [1–3], the transverse shear deformation is eliminated in thin shell analysis, such that at least C1 continuous shape functions are required within Galerkin methods. In static and dynamic simulation of structure, the conventional finite element methods [1,2] are one of the most popular approximation scheme, however the construction of C1 continuity is still a big challenge for cell-based finite element methods. In last three decades, the meshfree methods [1–3] equipped high order smoothed shape functions have attracted significant research attention, while the meshfree shape functions are established based upon a set of scattered nodes and the high order continuity of shape functions is easily fulfilled even with low order basis function. For thin shell analysis, this high order meshfree approximations can also alleviate the membrane locking caused by the mismatched approximation order of membrane strain and bending strain [1]. Moreover, in general, the nodal-based meshfree approximations can release the burden of mesh distortion and have the flexibility of local refinement. Due to these advantages, a wide variety meshfree methods are proposed and have been applied to many scientific or engineering fields. Among of them, moving least squares (MLS) and reproducing kernel (RK) meshfree approximations built their shape functions by enforcing the so-call consistency conditions, where the consistency conditions require that the corresponding approximations should exactly reproduce every functions spanned by basis functions, and this conditions usually serve as a basic requirement for the error convergence of resolved Galerkin solutions [1]. However, the high order smoothed meshfree shape functions accompany with the severely overlapping supports, which leads to a misalignment between numerical integration domains and supports of shape functions. As a result, the meshfree shape functions usually exhibit a piecewise rational nature in each integration domains, and it brings a serious difficulty to the accurate numerical integration in Galerkin weak forms [1].