### 2024-02-11 main.tex

\documentclass{elsarticle}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts,amsthm,bm}
\usepackage{hyperref}
\usepackage{booktabs,multirow}
\usepackage{lineno}
\linenumbers

\bibliographystyle{elsarticle-num}

\begin{document}
\begin{frontmatter}
    \title{A principle meshfree formulation for thin shells with naturally consistent enforcement of essential boundary conditions}
    \author[1]{Junchao Wu\corref{cor1}}
\end{frontmatter}
\include{introduction}
\include{kinematics}
\include{mix}
\include{boundary}
\include{examples}
\include{conclusion}
\appendix
\include{derivative}
\include{derivation}
\bibliography{references}
\end{document}