### 2024-02-11 examples.tex
temp

\section{Numerical examples}
\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{figures/patchtest_msh}
    \caption{Meshfree discretization for patch test}

\end{figure}

\begin{table}[h!]
\centering
\caption{Results of patch test for flat model}
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{Linear patch test} & \multicolumn{2}{c}{Quadratic patch test} \\ \cline{2-5}
 & $L_2$-Error & $H_e$-Error & $L_2$-Error & $H_e$-Error \\
    \midrule
    GI-Penalty & $4.45E-4$ & $1.35E-2$ & $2.01E-3$ & $1.63E-2$ \\
    GI-Nitsche & $4.51E-4$ & $1.42E-2$ & $1.22E-3$ & $1.68E-2$ \\
    RKGSI-Penalty & $3.64E-9$ & $6.77E-8$ & $4.54E-9$ & $6.57E-8$ \\
    RKGSI-Nitsche & $3.31E-12$ & $1.34E-11$ & $5.98E-12$ & $1.21E-11$ \\
    RKGSI-HR & $6.67E-13$ & $1.50E-11$ & $1.07E-12$ & $1.26E-11$ \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Results of patch test for curved model}
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{Linear patch test} & \multicolumn{2}{c}{Quadratic patch test} \\ \cline{2-5}
 & $L_2$-Error & $H_e$-Error & $L_2$-Error & $H_e$-Error \\
    \midrule
    GI-Penalty & $3.79E-4$ & $1.30E-2$ & $1.74E-3$ & $1.37E-2$ \\
    GI-Nitsche & $4.04E-4$ & $1.42E-2$ & $1.15E-3$ & $1.49E-2$ \\
    RKGSI-Penalty & $3.64E-9$ & $6.77E-8$ & $4.54E-9$ & $6.57E-8$ \\
    RKGSI-Nitsche & $3.31E-12$ & $1.34E-11$ & $5.98E-12$ & $1.21E-11$ \\
    RKGSI-HR & $4.28E-6$ & $1.30E-4$ & $9.69E-6$ & $1.30E-4$ \\
    \bottomrule
\end{tabular}
\end{table}