### 2024-02-14 conclusion.tex

\section{Conclusion}\label{conclusion}
An efficient and quasi-consistent meshfree thin shell formulation was presented to naturally enforce the essential boundary conditions. In this approach, the mixed formulation with Hu-Washizu principle weak form is employed, where the displacement is discretized by traditional meshfree shape functions, the strains and stresses can be expressed by reproducing kernel smoothed gradients and covariant smoothed gradients. The smoothed gradient naturally embed the first two order integration constraint, and has a quasi variational consistency for curved models in each integration cells. Owing to the Hu-Washizu variational principle, the essential boundary condition enforcement has a similar form with conventional Nitsche's method, both have the consistent term and stabilized term. Compared with Nitsche's method, the costly high order derivatives in Nitsche's consistent term have been replaced by smoothed gradients, which shows great computational speed due to the reproducing kernel gradient smoothing framework. Meanwhile, the stabilized term is naturally existed in Hu-Washizu weak form, and the artificial parameter needed in Nitsche's stabilized term has been vanished, which can automatically maintain the coercivity for stiffness matrix. Numerical results demonstrated that the proposed Hu-Washizu quasi-consistent meshfree thin shell formulation show great performance in terms of accuracy, efficiency and stability